Enforced Invariant Confluence

### 2020-01-13 doc/vldb_paper/figures/merge_reducible.tex
Renamed paper/ to vldb_paper/.
I'm working on the VLDB Journal submission now, so there's actually two
papers to deal with.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[scale=0.3]
    \draw[rounded corners] (0, 0) rectangle (16, 8);
    \draw[rounded corners] (1, 2.5) rectangle (11, 6.5);
    \draw[rounded corners] (6, 0.5) rectangle (15, 4.5);
    \draw[rounded corners] (7, 1.5) rectangle ++(7, 2);
    \node[inner sep=0pt, anchor=north] at (8, 7.5) {\small Invariant-Confluent};
    \node[inner sep=0pt, anchor=north] at (6, 6) {\small Invariant-Closed};
    \node[inner sep=0pt, anchor=south] at (10.5, 1.5) {\small Criteria (1) -- (4)};
    \node[inner sep=0pt, anchor=south] at (10.5, 0.5) {\small Merge-Reducible};
  \end{tikzpicture}
  \caption{%
    The relationship between invariant-closure, merge-reducibility, criteria
    (1) -- (4) from \thmref{LatticeProperty}, and invariant-confluence.
  }\figlabel{IclosureVsReducible}
\end{figure}

>>>>

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[scale=0.3]
    \draw[rounded corners] (0, 0) rectangle (16, 8);
    \draw[rounded corners] (1, 2.5) rectangle (11, 6.5);
    \draw[rounded corners] (6, 0.5) rectangle (15, 4.5);
    \draw[rounded corners] (7, 1.5) rectangle ++(7, 2);
    \node[inner sep=0pt, anchor=north] at (8, 7.5) {\scriptsize Invariant Confluent};
    \node[inner sep=0pt, anchor=north] at (6, 6) {\scriptsize Invariant Closed};
    \node[inner sep=0pt, anchor=south] at (10.5, 1.5) {\scriptsize Criteria (1) -- (4)};
    \node[inner sep=0pt, anchor=south] at (10.5, 0.5) {\scriptsize Merge Reducible};
  \end{tikzpicture}
  \caption{%
    The relationship between \invariantclosure{}, \mergereducibility{},
    criteria (1) -- (4) from \thmref{LatticeProperty}, and
    \invariantconfluence{}.
  }\figlabel{IclosureVsReducible}
\end{figure}

### 2020-01-13 doc/vldb_paper/sections/related_work.tex
Renamed paper/ to vldb_paper/.
I'm working on the VLDB Journal submission now, so there's actually two
papers to deal with.

\textbf{Token Based Invariant Confluence.}
In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token based
consistency model that generalizes a family of consistency models including
causal consistency, sequential consistency, and RedBlue consistency. An
application designer defines a set of tokens and specifies which pairs of
tokens conflict, and transactions acquire some subset of the tokens when they
execute. This allows the application designer to specify which transactions
conflict with one another.  Gotsman et al.\ develop sufficient conditions to
determine whether a given token scheme is sufficient to guarantee that a global
invariant is never broken. This model of consistency is very similar to
segmented \invariantconfluence{}. The token based approach allows users to
specify certain conflicts that are not possible with segmented
\invariantconfluence{}. A segmentation only allows transactions within a
segment to acquire a single self-conflicting lock. However, segmented
\invariantconfluence{} also introduces the notion of invariant segmentation,
which cannot be emulated with the token based approach. For example, it is
difficult to emulate escrow transactions with the token based approach.

@@

\textbf{CRDTs.}
CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
semilattices with inflationary update methods. Due to their algebraic
properties, CRDTs can be replicated with strong eventual consistency without
the need for any coordination. Our definition of distributed objects and our
\invariantconfluence{} system model are inspired directly by the corresponding
definitions and system models in~\cite{shapiro2011conflict}.
%
CRDTs are eventually consistent but may not preserve invariants. Conversely,
\invariantconfluent{} objects preserve invariants but may not be eventually
consistent. Thus, it is natural (though not necessary) to use CRDTs as
distributed objects. If a CRDT is determined to be \invariantconfluent{} with
respect to a particular invariant and set of transactions, then it achieves a
combination of strong eventual consistency and invariant preservation. Any CRDT
(e.g., counters, sets, graphs, sequences) can be used for this purpose.
%
Finally, our criteria in \thmref{LatticeProperty} also borrow ideas from CRDTs,
exploiting the nice algebraic properties of semilattices.

\textbf{CALM Theorem.}
Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
formalism, Dedalus~\cite{alvaro2011dedalus, alvaro2013declarative}, are
declarative Datalog-based programming languages that are designed to program
distributed systems. The accompanying CALM
theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
and only if a program can be written in the monotone fragment of these
languages, then there exists a consistent, coordination-free implementation of
the program.  The CALM theorem provides guarantees about the consistency of
program outputs. It does not directly capture our notions of transactions or
invariant maintenance during program execution.  Moreover, Bloom and Dedalus
are general purpose programming languages that can be used to implement a
variety of distributed systems that are outside of the scope of
\invariantconfluence{}.

>>>>

\textbf{Token Based Invariant Confluence.}
In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token based
consistency model that generalizes a family of consistency models including
causal consistency, sequential consistency, and RedBlue consistency. An
application designer defines a set of tokens and specifies which pairs of
tokens conflict, and transactions acquire some subset of the tokens when they
execute. This allows the application designer to specify which transactions
conflict with one another.  Gotsman et al.\ develop sufficient conditions to
determine whether a given token scheme is sufficient to guarantee that a global
invariant is never broken. The token based approach allows users to
specify certain conflicts that are not possible with segmented
\invariantconfluence{} because a segmentation only allows transactions within a
segment to acquire a single self-conflicting lock. However, segmented
\invariantconfluence{} also introduces the notion of invariant segmentation,
which cannot be emulated with the token based approach. For example, it is
difficult to emulate escrow transactions with the token based approach.

@@

\textbf{CRDTs.}
CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
semilattices with inflationary update methods. Due to their algebraic
properties, CRDTs can be replicated with strong eventual consistency without
the need for any coordination. Our definition of distributed objects and our
\invariantconfluence{} system model are inspired directly by the corresponding
definitions and system models in~\cite{shapiro2011conflict}.
%
CRDTs are eventually consistent but may not preserve invariants. Conversely,
\invariantconfluent{} objects preserve invariants but may not be eventually
consistent. Thus, it is natural (though not necessary) to use CRDTs as
distributed objects. If a CRDT is determined to be \invariantconfluent{} with
respect to a particular invariant and set of transactions, then it achieves a
combination of strong eventual consistency and invariant preservation. Any CRDT
(e.g., counters, sets, graphs, sequences) can be used for this purpose.
%
Finally, our criteria in \thmref{LatticeProperty} also borrow ideas from CRDTs,
exploiting the algebraic properties of semilattices.

\textbf{CALM Theorem.}
Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
formalism, Dedalus~\cite{alvaro2011dedalus, alvaro2013declarative}, are
declarative Datalog-based programming languages that are designed to program
distributed systems. The accompanying CALM
theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
and only if a program can be written in the monotone fragment of these
languages, then there exists a consistent, coordination-free implementation of
the program.  The CALM theorem provides guarantees about the consistency of
program outputs. It does not directly capture our notions of transactions or
invariant maintenance during program execution.  Moreover, Bloom and Dedalus
are general-purpose programming languages that can be used to implement a
variety of distributed systems that are outside of the scope of
\invariantconfluence{}.
